Short_note_captured_phenotype

### 2025-02-03 Manuscript_data_analyses.qmd
Last formatting changes

library(texreg) #export table in Latex format
library(kableExtra)

@@

Declaration_Bycatch[TARGETED_TAXA == "MZZ - Osteichthyes", TARGETED_TAXA:=NA]
OBSMER_Bycatch[TARGETED_TAXA == "MZZ - Osteichthyes", TARGETED_TAXA:=NA] # we considered the declarations of Osteichtyes to be equivalent to DECLARATIONS OF an unknown taxon, and therefore considered these values equivalent to NA for our analyses

# Merging data in common between declaration and observation datasets:

@@

# Merging data in common between declaration and observation datasets, by only considering explanatory variables:

@@

# Maps of corrected abundance:

@@

allgeo=allgeo[allgeo$offname %in% c("27.4.c", "27.7.d", "27.7.e", "27.7.f", "27.7.g", "27.7.h", "27.7.j", "27.7.k", "27.8.a", "27.8.b", "27.8.c", "27.8.d"),] #selection of only ices division along the French coast

@@

We have separated the analysis of stranding data from the declaration and observation data because they may be partially redundant (with no way to verify this), and because their collection is fundamentally different (only some of the captured dolphins strand, whereas theoretically we should detect all captures through observation and reporting).

For each data sets, we first described the distribution of studied traits and compared the trait values of by-caught and stranded individuals with key population values reported in the literature for each of the species studied (for both sexes: body length at maturity, asymptotic adult body length). We then examined whether different phenotypic sensitivity profiles emerged in space or time by implementing a generalised linear model for each of the traits studied (linear model for body length, logistic regression for sex) as a function of year, trimester and ICES division of by-catch.

@@

# Creation of datasets per species:

@@

# key population values reported in the literature for each of the species studied (for both sexes: body length at maturity, asymptotic adult body length)

@@

# some years appears quite rarely and the number of categories associated with the variable is high

@@

# only model on size are performed because this measure is systematically performed for stranded individuals (contrary to weight) and more reliable in this specific case

@@

# only model on size are performed because this measure is systematically performed for stranded individuals (contrary to weight) and more reliable in this specific case

@@

### Table combining pairwise contrast comparison test for each significant variable either in models studying sex distribution (S3) or size distribution (S4):

@@

# The distribution of the number of captures per ICES division is highly variable, with some underrepresented divisions. Due to rare categories and the fact that the sample size of recorded captures is too small to support analyses with categorical variables described by a high number of classes, we decided to group years into four different groups.

@@

table(OBSMER_Bycatch[!is.na(SEX) & SPECIES == "Delphinus delphis"]$ICES_DIV_grouped)
fisher.test(OBSMER_Bycatch[!is.na(SEX) & SPECIES == "Delphinus delphis"]$SEX, OBSMER_Bycatch[!is.na(SEX) & SPECIES == "Delphinus delphis"]$ICES_DIV_grouped)
table(OBSMER_Bycatch[!is.na(SEX) & SPECIES == "Delphinus delphis"]$SEX, OBSMER_Bycatch[!is.na(SEX) & SPECIES == "Delphinus delphis"]$ICES_DIV_grouped) # more female capture north of the bay of Biscay (not significant)

@@

# Prior to spatio-temporal test we checked whether body mass variation depended on the data sources:

@@

table(OBSMER_Bycatch[SPECIES == "Delphinus delphis"]$SEA_STATE)
table(OBSMER_Bycatch[SPECIES == "Phocoena phocoena"]$SEA_STATE)
# as their is some sea state categories with too little data (highest and lowest sea state measured) we grouped the most extreme conditions together for analyses:

@@

table(OBSMER_Bycatch[SPECIES == "Delphinus delphis",]$SEA_STATE_grouped, OBSMER_Bycatch[SPECIES == "Delphinus delphis",]$TRIMESTER)
table(OBSMER_Bycatch[SPECIES == "Delphinus delphis",]$SEA_STATE_grouped, OBSMER_Bycatch[SPECIES == "Delphinus delphis",]$ICES_DIV_grouped)
chisq.test(OBSMER_Bycatch[SPECIES == "Delphinus delphis",]$SEA_STATE_grouped, OBSMER_Bycatch[SPECIES == "Delphinus delphis",]$TRIMESTER)
chisq.test(OBSMER_Bycatch[SPECIES == "Delphinus delphis",]$SEA_STATE_grouped, OBSMER_Bycatch[SPECIES == "Delphinus delphis",]$ICES_DIV_grouped)
# For the interpretation of sea state effects: trimester and spatial areas are highly correlated with sea conditions: calm sea conditions were mostly reported in the Bay of Biscay and rough sea conditions mainly in the Channel Sea; these two latter conditions were also mostly reported in winter

@@

We then explored the susceptibility of individuals to be captured as a function of fishing technique. These analyses initially considered only the risk of being captured (see first graphical outputs, this part was scanned quickly as it had already been thoroughly studied) and how this susceptibility varies with individual traits (i.e. phenotypic vulnerabilities to fishing technique, see subsequent regression models).

@@

# We first considered only the risk of being captured as a function of fishing technique. This part was scanned quickly as it had already been thoroughly studied by other researches. We only produced graphical outputs illustrating the relationships between fishing behaviours and the bumber of captured individuals.

@@

@@

We then studied how traits of by-caught individuals varied with fishing techniques used so as to identify potential phenotypic vulnerabilities. Three phenotypic traits were studied: the by-caught individuals' body mass (using both declaration and observation data), body length and sex (using observation data only).

@@

# We began to remove too rare categories (N<6) from regression analyses (while having removed all missing data with missing values for at least one of the considered explanatory variables):

@@

table(Decl_Obs_Merged_Bycatch[SPECIES == "Phocoena phocoena" & !is.na(TARGETED_ISSCAAP) & !is.na(MESH_SIZE) & !is.na(FISHING_GEAR_grouped) & TARGETED_ISSCAAP %in% c("31 - Pleuronectiformes", "32 - Gadiformes", "34 - Miscellaneous demersal fishes")]$FISHING_GEAR_grouped)
# in this specific case there is rare categories in both categorical variables: removing rare categories in targeted ISSCAAP groups categories lead to the presence of too rare categories in used fishing gears. We therefore cannot use both variables in the model. In this particular case we chose to favor the use of used fishing gear in explanatory variables.

@@

apply(OBSMER_Bycatch, 2, function(x) table(is.na(x)))
# some variable present too many missing values to be included in analyses (as gear speed only available for trawls)

@@

table(OBSMER_Bycatch[SPECIES == "Phocoena phocoena" & !is.na(BODY_LENGTH) & !is.na(MOST_FISHED_ISSCAAP) & !is.na(MESH_SIZE) & !is.na(FISHING_GEAR_grouped) & MOST_FISHED_ISSCAAP %in% c("31 - Pleuronectiformes", "34 - Miscellaneous demersal fishes")]$FISHING_GEAR_grouped)
# in this specific case there is rare categories in both categorical variables: removing rare categories in targeted ISSCAAP groups categories lead to the presence of too rare categories in used fishing gears. We therefore cannot use both variables in the model. In this particular case we chose to favor the use of used fishing gear in explanatory variables.

@@

@@

############ Modifier legende axe abscisse pour simplifier interpretation ?????

#reordering variables for plots:

# OBSMER_size_del$ISSCAAP=factor(OBSMER_size_del$ISSCAAP, levels = c("31","38","32","36","37","33","34"), labels=c("Pleuronectiformes\n(mainly soles)", "Sharks, rays", "Gadiformes\n(mainly hake and gadoids)", "Tunas, bonitos", "Miscellaneous pelagic fishes\n(mainly mackerells)", "Miscellaneous coastal fishes\n(mainly sea bass)", "Miscellaneous demersal fishes\n(mainly monkfishes or St Pierre)")) # 31:poisson plats, 32:merlus et gadidés, 33: mainly bass, 34: démerseaux (baudroies, saint pierre), 36: thon, bonite, 37: maqueraux, chinchard, 38: requins, raies

@@

### Global table (Table S6) of vulnerabilities regression models (ANOVA, type II):

@@

### Table (S7) combining pairwise contrast comparison test for each significant variable in vulnerabilities regression models:

>>>>

library(texreg) 
library(kableExtra) #export table in Latex format

@@

Declaration_Bycatch[TARGETED_TAXA == "MZZ - Osteichthyes", TARGETED_TAXA:=NA]
OBSMER_Bycatch[TARGETED_TAXA == "MZZ - Osteichthyes", TARGETED_TAXA:=NA] # we considered the declarations of Osteichtyes to be equivalent to declarations of an unknown taxon, and therefore considered these values equivalent to NA for our analyses

# Merging data shared by declaration and observation datasets:

@@

# Merging data shared by declaration and observation datasets, but only considering explanatory variables:

@@

# Map of corrected abundance:

@@

allgeo=allgeo[allgeo$offname %in% c("27.4.c", "27.7.d", "27.7.e", "27.7.f", "27.7.g", "27.7.h", "27.7.j", "27.7.k", "27.8.a", "27.8.b", "27.8.c", "27.8.d"),] #selection of sampled ices divisions

@@

We have separated the analysis of stranding data from the declaration and observation data because they may be partially redundant (with no way to verify this), and because their collection is fundamentally different (only some of the captured dolphins actually strand, whereas theoretically we should detect all captures through observation and reporting).

For each dataset, we first described the distribution of studied traits and compared the trait values of by-caught and stranded individuals with key population values reported in the literature for each of the species studied (for both sexes: body length at maturity, asymptotic adult body length). We then examined whether different phenotypic sensitivity profiles emerged in space or time by implementing a generalised linear model for each of the traits studied (linear model for body length, logistic regression for sex) as a function of year, trimester and ICES division of by-catch.

@@

# Creation of species-specific datasets:

@@

# biblio: key population values reported in the literature for each of the species studied (for both sexes: body length at maturity, asymptotic adult body length)

@@

# some years appear quite rarely and the number of categories associated with the variable is high

@@

# only model on size are performed because this measure is systematically performed for stranded individuals (contrary to weight) and more reliable in our specific case (body mass values are guess estimations)

@@

# only model on size are performed because this measure is systematically performed for stranded individuals (contrary to weight) and more reliable in our specific case

@@

### Table combining pairwise contrast comparison tests for each significant variable either in models studying sex distribution (S3) or size distribution (S4):

@@

# The distribution of the number of captures per ICES division is highly variable, with some underrepresented divisions. Due to rare categories and the fact that the sample size of recorded captures is too small to support analyses with categorical variables described by a high number of classes, we decided to group ICES divisions into four different groups.

@@

table(OBSMER_Bycatch[!is.na(SEX) & SPECIES == "Delphinus delphis"]$ICES_DIV_grouped)
fisher.test(OBSMER_Bycatch[!is.na(SEX) & SPECIES == "Delphinus delphis"]$SEX, OBSMER_Bycatch[!is.na(SEX) & SPECIES == "Delphinus delphis"]$ICES_DIV_grouped)
table(OBSMER_Bycatch[!is.na(SEX) & SPECIES == "Delphinus delphis"]$SEX, OBSMER_Bycatch[!is.na(SEX) & SPECIES == "Delphinus delphis"]$ICES_DIV_grouped) # more female captured north of the bay of Biscay (not significant)

@@

# Prior to spatio-temporal tests we checked whether body mass variation depended on the data sources:

@@

table(OBSMER_Bycatch[SPECIES == "Delphinus delphis"]$SEA_STATE)
table(OBSMER_Bycatch[SPECIES == "Phocoena phocoena"]$SEA_STATE)
# as there is some sea state categories with too little data (highest and lowest sea state measured) we grouped the most extreme conditions together for analyses:

@@

table(OBSMER_Bycatch[SPECIES == "Delphinus delphis",]$SEA_STATE_grouped, OBSMER_Bycatch[SPECIES == "Delphinus delphis",]$TRIMESTER)
table(OBSMER_Bycatch[SPECIES == "Delphinus delphis",]$SEA_STATE_grouped, OBSMER_Bycatch[SPECIES == "Delphinus delphis",]$ICES_DIV_grouped)
chisq.test(OBSMER_Bycatch[SPECIES == "Delphinus delphis",]$SEA_STATE_grouped, OBSMER_Bycatch[SPECIES == "Delphinus delphis",]$TRIMESTER)
chisq.test(OBSMER_Bycatch[SPECIES == "Delphinus delphis",]$SEA_STATE_grouped, OBSMER_Bycatch[SPECIES == "Delphinus delphis",]$ICES_DIV_grouped)
# For the interpretation of sea state effects: trimester and spatial areas are highly correlated with sea conditions; calm sea conditions were mostly reported in the Bay of Biscay and rough sea conditions mainly in the Channel Sea, these two latter conditions were also mostly reported in winter

@@

We then explored the susceptibility of individuals to be captured as a function of fishing techniques. These analyses initially considered only the risk of being captured (see first graphical outputs, this part was scanned quickly as it had already been thoroughly studied) and how this susceptibility varies with individual traits (i.e. phenotypic vulnerabilities to fishing technique, see subsequent regression models).

@@

# We first considered only the risk of being captured as a function of fishing techniques. This part was scanned quickly as it had already been thoroughly studied by other researches. We only produced graphical outputs illustrating the relationships between fishing behaviours and the number of captured individuals.

@@

# To better illustrate the distribution of mostly fished taxa associated with by-catch, we defined the proportion of most fished species or group of species mainly associated with mostly fished ISSCAAP groups (only for ISSCAAP groups with more than 6 by-catch observations, as other groups will not be considered for analyses because too rare). The proportion were computed from the above tables and the total number of data available for the most fished associated ISSCAAP group.

@@

We then studied how traits of by-caught individuals varied with used fishing techniques, so as to identify potential phenotypic vulnerabilities. Three phenotypic traits were studied: the by-caught individuals' body mass (using both declaration and observation data), body length and sex (using observation data only).

@@

# We began to remove too rare categories (N<6) from regression analyses (while having removed all by-catch data with missing values for at least one of the considered explanatory variables):

@@

table(Decl_Obs_Merged_Bycatch[SPECIES == "Phocoena phocoena" & !is.na(TARGETED_ISSCAAP) & !is.na(MESH_SIZE) & !is.na(FISHING_GEAR_grouped) & TARGETED_ISSCAAP %in% c("31 - Pleuronectiformes", "32 - Gadiformes", "34 - Miscellaneous demersal fishes")]$FISHING_GEAR_grouped)
# in this specific case there is rare categories in both categorical variables: removing rare categories in targeted ISSCAAP group categories lead to the presence of too rare categories in used fishing gears. We therefore cannot use both variables in the model. In this particular case, we chose to favor the use of used fishing gear in explanatory variables.

@@

apply(OBSMER_Bycatch, 2, function(x) table(is.na(x)))
# some variables present too many missing values to be included in analyses (as gear speed only available for trawls)

@@

table(OBSMER_Bycatch[SPECIES == "Phocoena phocoena" & !is.na(BODY_LENGTH) & !is.na(MOST_FISHED_ISSCAAP) & !is.na(MESH_SIZE) & !is.na(FISHING_GEAR_grouped) & MOST_FISHED_ISSCAAP %in% c("31 - Pleuronectiformes", "34 - Miscellaneous demersal fishes")]$FISHING_GEAR_grouped)
# in this specific case there are rare categories in both categorical variables: removing rare categories in targeted ISSCAAP groups categories lead to the presence of too rare categories in used fishing gears. We therefore cannot use both variables in the model. In this particular case we chose to favor the use of used fishing gear in explanatory variables.

@@

# we then checked for collinearity between explanatory variables (and for the inclusion of potential random effects, only when there are significant non-parametric tests of spatio-temporal (and source) effects on the phenotype of by-catch individuals):

@@

@@

### Global table (Table S6) of phenotypic vulnerabilities regression models (ANOVA, type II):

@@

### Table (S7) combining pairwise contrast comparison tests for each significant variable in vulnerabilities regression models:

### 2025-02-03 Manuscript_data_curation.qmd
Last formatting changes

library(stringr)# handling string format in R 
library(data.table) #large dataset gestion 
library(sf) #spatial treatments
library(lubridate) #handling date data

sumna <- function(x) {
    ret <- sum(x, na.rm = TRUE)
    if(ret == 0 && all(is.na(x))) {NA_real_} else {ret}
}

@@

# we only retrieved here fishing activity information that will use latter on for complement our original dataset

for (i in 2000:2023) {  #the year 2023 is included to take into account the marees beginning in 2022 and ending in 2023 (including not negligible amount of data about the end of 2022)

@@

SACROIS[,NAVS_COD_YEAR:=paste(SACROIS[,NAVS_COD], str_sub(SACROIS[,DATE_SEQ],7,10), sep="_")]  #produce a unique ID per vessel ad per year

#minimal requirement for being able to run the previous section: 16GB of RAM

#important notice: the last studied year (here 2022) does not take into account overlapping marees (between this year and the following one): have to be corrected by taking the following year into account during computation and then removing it at the end !!!! Important to do  because represent a non-negligible amount of data at the end of the year

@@

The purpose of this code is to format the data so that it can be analysed later. Three datasets were used: the fisheries observer data (OBSMER data) and the French data on reported fishing activities (SACROIS data, which also includes the skipper declaration of by-catches). We prepared data to retain only variables used for analysis, time and areas of interest, while removing potentially error-prone data.

@@

DECL_BYC=DECL_BYC[ESP_LIB_FAO %in% c("Dauphin à flancs blancs Atlan.", "Dauphin bleu et blanc","Dauphin commun","Dauphin commun a bec large","Dauphin tacheté de l'Atlantiq.","Dauphins nca","Grand dauphin", "Marsouin commun")] # we only retained by-catch data on small cetacean

@@

# there is no individual traits measurements for declaration of multiple capture, we then discarded such declarations from analyses:

@@

# We discarded all data with individuals weighing less than 3 kg or more than 300 kg (range of values outside the weight ranges existing in natural conditions)

@@

# We discarded all data with individuals weighing more than 150 kg (range of values outside the weight ranges existing in natural conditions)

@@

# Selection of areas of interest and retrieval of ICES division associated with location points of strandings:

@@

# Removal of data without any phenotypic measure:

@@

OBSMER_BYC=merge(OBSMER_BYC, strat[,.(ID_VRAC,ID_OP)], by="ID_VRAC", all.x = T, all.y=F) #retrieval of information on fishing activity during fishing operation

@@

OBSMER_BYC=merge(OBSMER_BYC, marees, by="ID_MAREE", all.x = T, all.y=F)

@@

#renaming confusing variables:

@@

capt=capt[!(str_detect(ESPECE, "Macrodéchets|Inerte|Cetorhinus maximus|Animaux|Halichoerus grypus|Uria aalge"))] #removing catch information on waste, other megafauna bycatch (not commercialized species), or undetermined taxa

# retrieving catch weight for each captured taxa

@@

OBSMER_BYC=rbind(OBSMER_BYC[NOMBRE.y==1 | is.na(NOMBRE.y)], OBSMER_BYC[NOMBRE.y>1][rep(sequence(nrow(OBSMER_BYC[NOMBRE.y>1])), OBSMER_BYC[NOMBRE.y>1, NOMBRE.y]),]) # for the number indicated in measurements, number superior to one correspond to dolphin having the same measurements (we therefore repeated these lines)

@@

for (i in 1:length(u)) {
    if (u[i]==1 | (u[i]>1 & cat[i]==1)) {
      for (j in 1:u[i]) {
        OBSMER_BYC[ID_CAPTURE==unique(OBSMER_BYC$ID_CAPTURE)[i],][which(!duplicated(CRITERE))+(j-1),]$ID_IND=
        paste(OBSMER_BYC[ID_CAPTURE==unique(OBSMER_BYC$ID_CAPTURE)[i],]$ID_CAPTURE[1],j, sep="_")
      }
    }
    else {
      for (j in 1:length(OBSMER_BYC[ID_CAPTURE==unique(OBSMER_BYC$ID_CAPTURE)[i], ID_CAPTURE])) {
        OBSMER_BYC[ID_CAPTURE==unique(OBSMER_BYC$ID_CAPTURE)[i],][j,]$ID_IND=
          paste(OBSMER_BYC[ID_CAPTURE==unique(OBSMER_BYC$ID_CAPTURE)[i],]$ID_CAPTURE[1],j, sep="_")
      } # individuals measured in groups, with the same capture ID, and with multiple measures: considering the different measures as different individuals as we are not able to reliably assigned the different measurements to each individual
    }
}

@@

OBSMER_BYC[POIDS_REF.x %in% c("0.0", "10.0", "15.0", "999.0"), POIDS_REF.x:=""] # absurd values (also when compared to size measurements, which are much more reliable as weight are guess estimates)
OBSMER_BYC[ESPECE=="Phocoena phocoena" & POIDS_REF.x=="150.0", POIDS_REF.x:=""] # discarding all data with individuals weighing more than 150 kg (range of values outside the weight ranges existing in natural conditions)

@@

# removal of 4 measures of circonference completely aberrant (<30cm for weight >100kg, or >180cm...)
OBSMER_BYC[CRITERE=="Circonférence (C)" & as.numeric(VALEUR)<35, VALEUR:=""]
OBSMER_BYC[CRITERE=="Circonférence (C)" & as.numeric(VALEUR)>170, VALEUR:=""]

@@

@@

# trip IDs from declaration and observation datasets are not attributed the same way

@@

for (i in 1:length(OBSMER_BYC_simp$NAVS_COD_YEAR)) {
  OBSMER_BYC_simp[i, MAREE_ID:=c(names(which.max(table(SACROIS[NAVS_COD_YEAR==OBSMER_BYC_simp[i, NAVS_COD_YEAR] & LIEU_COD_RET_SACROIS==OBSMER_BYC_simp[i, PORT_DEBARQUEMENT_CODE] & dmy_hms(DATE_SEQ)>=dmy_hms(OBSMER_BYC_simp[i, DATE_DEBUT_MAREE]) & dmy_hms(DATE_SEQ)<=dmy_hms(OBSMER_BYC_simp[i, DATE_FIN_MAREE]) & !duplicated(SEQ_ID), MAREE_ID]))), NA_integer_)[1]] #in case of overlapping trips: select the ones with the highest number of sequence included in target trip
} # retrieving the most probable declaration fishing trip ID (on the basis of landing harbour and fishing trip timing) for each vessels of potential overlap between declaration and observation data

@@

# Declaration data:

>>>>

library(stringr)# handling string format in R 
library(data.table) # large dataset gestion 
library(sf) # spatial treatments
library(lubridate) # handling date data

sumna <- function(x) {
    ret <- sum(x, na.rm = TRUE)
    if(ret == 0 && all(is.na(x))) {NA_real_} else {ret}
} # sum function returning NA when summing NA values only

@@

# we only retrieved here fishing activity information that will be used latter on, to complement information from our initial datasets

for (i in 2000:2023) {  #the year 2023 is included to take into account the fishing trip beginning in 2022 and ending in 2023 (including not negligible amount of data about the end of 2022)

@@

SACROIS[,NAVS_COD_YEAR:=paste(SACROIS[,NAVS_COD], str_sub(SACROIS[,DATE_SEQ],7,10), sep="_")]  #produce a unique ID per vessel and per year

#minimal requirement for being able to run the previous section of code: 16GB of RAM

@@

The purpose of this code is to format the data so that it can be analysed later. Three datasets were used: the fisheries observer data (OBSMER data), the French data on reported fishing activities (SACROIS data, which also includes the skipper declaration of by-catches), and French stranding records ('Réseau national d'échouages'). We prepared data to retain only variables used for analysis, time periods and areas of interest, while removing potentially error-prone data.

@@

DECL_BYC=DECL_BYC[ESP_LIB_FAO %in% c("Dauphin à flancs blancs Atlan.", "Dauphin bleu et blanc","Dauphin commun","Dauphin commun a bec large","Dauphin tacheté de l'Atlantiq.","Dauphins nca","Grand dauphin", "Marsouin commun")] # We only retained by-catch data on small cetacean

@@

# there are no individual traits measurements for declaration of multiple captures, we then discarded such declarations from analyses:

@@

# We discarded all data from individuals weighing less than 3 kg or more than 300 kg (range of values outside the weight ranges existing in natural conditions)

@@

# We discarded all data from individuals weighing more than 150 kg (range of values outside the weight ranges existing in natural conditions)

@@

# Selection of areas of interest and retrieval of ICES divisions associated with location points of strandings:

@@

# Removal of data without any phenotypic measurements:

@@

OBSMER_BYC=merge(OBSMER_BYC, strat[,.(ID_VRAC,ID_OP)], by="ID_VRAC", all.x = T, all.y=F) # retrieval of information on fishing activity during fishing operation

@@

OBSMER_BYC=merge(OBSMER_BYC, marees, by="ID_MAREE", all.x = T, all.y=F) #retrieval of information on fishing trip

@@

# renaming confusing variables:

@@

capt=capt[!(str_detect(ESPECE, "Macrodéchets|Inerte|Cetorhinus maximus|Animaux|Halichoerus grypus|Uria aalge"))] # removing catch information on waste, other megafauna bycatch (not commercialized species), or undetermined taxa

# retrieving catch weight for each captured taxa:

@@

OBSMER_BYC=rbind(OBSMER_BYC[NOMBRE.y==1 | is.na(NOMBRE.y)], OBSMER_BYC[NOMBRE.y>1][rep(sequence(nrow(OBSMER_BYC[NOMBRE.y>1])), OBSMER_BYC[NOMBRE.y>1, NOMBRE.y]),]) # for the number of individuals indicated in measurements, number superior to one correspond to dolphin having the same measurements (we therefore repeated these lines)

@@

for (i in 1:length(u)) {
    if (u[i]==1 | (u[i]>1 & cat[i]==1)) {
      for (j in 1:u[i]) {
        OBSMER_BYC[ID_CAPTURE==unique(OBSMER_BYC$ID_CAPTURE)[i],][which(!duplicated(CRITERE))+(j-1),]$ID_IND=
        paste(OBSMER_BYC[ID_CAPTURE==unique(OBSMER_BYC$ID_CAPTURE)[i],]$ID_CAPTURE[1],j, sep="_")
      }
    }
    else {
      for (j in 1:length(OBSMER_BYC[ID_CAPTURE==unique(OBSMER_BYC$ID_CAPTURE)[i], ID_CAPTURE])) {
        OBSMER_BYC[ID_CAPTURE==unique(OBSMER_BYC$ID_CAPTURE)[i],][j,]$ID_IND=
          paste(OBSMER_BYC[ID_CAPTURE==unique(OBSMER_BYC$ID_CAPTURE)[i],]$ID_CAPTURE[1],j, sep="_")
      } # individuals measured in groups, with the same capture ID, and with multiple measures: considering the different measures as different individuals as we are not able to reliably assign the different measurements to each individual
    }
}

@@

OBSMER_BYC[POIDS_REF.x %in% c("0.0", "10.0", "15.0", "999.0"), POIDS_REF.x:=""] # absurd values (also when compared to size measurements, which are much more reliable as weight are guess estimates)
OBSMER_BYC[ESPECE=="Phocoena phocoena" & POIDS_REF.x=="150.0", POIDS_REF.x:=""] # discarding all data from individuals weighing more than 150 kg (range of values outside the weight ranges existing in natural conditions)

@@

# removal of 4 measures of circumference completely aberrant (<30cm for weight >100kg, or >180cm...)
OBSMER_BYC[CRITERE=="Circonférence (C)" & as.numeric(VALEUR)<35, VALEUR:=""]
OBSMER_BYC[CRITERE=="Circonférence (C)" & as.numeric(VALEUR)>170, VALEUR:=""]

@@

# Important note: trip IDs from declaration and observation datasets are not attributed the same way

@@

@@

for (i in 1:length(OBSMER_BYC_simp$NAVS_COD_YEAR)) {
  OBSMER_BYC_simp[i, MAREE_ID:=c(names(which.max(table(SACROIS[NAVS_COD_YEAR==OBSMER_BYC_simp[i, NAVS_COD_YEAR] & LIEU_COD_RET_SACROIS==OBSMER_BYC_simp[i, PORT_DEBARQUEMENT_CODE] & dmy_hms(DATE_SEQ)>=dmy_hms(OBSMER_BYC_simp[i, DATE_DEBUT_MAREE]) & dmy_hms(DATE_SEQ)<=dmy_hms(OBSMER_BYC_simp[i, DATE_FIN_MAREE]) & !duplicated(SEQ_ID), MAREE_ID]))), NA_integer_)[1]] #in case of overlapping trips: select the ones with the highest number of sequence included in target trip
} # retrieving the most probable declaration fishing trip ID (on the basis of landing harbour and fishing trip timing) for each vessel with potential overlap between declaration and observation data

@@

# Declaration data (from SACROIS database):
